# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r loaddata, echo=TRUE, result="asis"}  
setwd("e://courses//jh-dataexplore//ws_research//RepData_PeerAssessment1")  
data <- read.csv("activity.csv")  
# sum up the steps per date, and get the steps  
step <- aggregate(steps ~ date, data, sum)$steps  
step 
```


```{r Histogram-stepday, fig.width=7, fig.height=6}
hist(step)  
```


## What is mean total number of steps taken per day?
mean and median is given below
```{r echo=TRUE}
mean(step)
median(step)
```


## What is the average daily activity pattern?
```{r echo=TRUE}
library(plyr)
stepdate <- ddply(data, .(interval), summarize, steps = mean(steps, na.rm=T))
```

```{r timeSeris-dailyactivity, fig.width=7, fig.height=6}
plot(stepdate$interval, stepdate$steps,type = "l", ylab="avg no of step", xlab="5-min interval", main="Time Series Plot")
```

interval that contains the maximum number of step
```{r echo=TRUE}
stepdate[stepdate$steps==max(stepdate$steps),]$interval
```


## Imputing missing values

```{r echo=TRUE}
print( "number of rows with missing value")
nrow(data[is.na(data),])
```


Strategy for filling in all of the missing values in the dataset:
calcuate the median of each 5-min interval for the whole day.  and use this to
fill in the missing value
```{r echo=TRUE}
data.nomiss <- data
intMedian <- ddply(data.nomiss, .(interval), summarize, steps = median(steps, na.rm=T))
data.nomiss[is.na(data.nomiss)]=intMedian$steps
stepNomiss <- aggregate(steps ~ date, data.nomiss, sum)$steps  
```

### histogram with missing data estimated
```{r histogram-adjustedData, fig.width=7, fig.height=6}
hist(stepNomiss)  
```


### mean and median of dataset with missing data estimated?
mean and median is given below
```{r echo=TRUE}
mean(stepNomiss);median(stepNomiss)
print("This differ from values of the dataset that has missing values as follows")
mean(step);median(step)
```


## Are there differences in activity patterns between weekdays and weekends?

method to create the appropriate dataframe required for plotting
(1) add a column "day" to the dataframe default the value to "weekday"
(2) change the value of this to weekend, if the date is either Saturday or Sunday
(3) get the mean of each interval of the day and group it by weekday and weekend into another dataframe
(4) use this dataframe (obtained in step3 ) for plotting
```{r echo=TRUE}
data.nomiss$day = "weekday"
data.nomiss[weekdays(as.Date(data.nomiss$date, "%Y-%m-%d")) == "Sunday" | weekdays(as.Date(data.nomiss$date, "%Y-%m-%d")) == "Saturday", ]$day="weekend"
stepdate.nomiss <- ddply(data.nomiss, .(interval, day), summarize, steps = mean(steps, na.rm=T))
```

use lattice to plot the graph
```{r TimeSeries-Weekend, fig.width=9, fig.height=6}
library(lattice)
xyplot(steps ~ interval |day, data=stepdate.nomiss, type ="l", layout=c(1,2), ylab="Number of steps")
```